---
title: "Differential abundance"
author: "Fernando Lucas Ruiz (fernando.lucas@um.es)"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    highlight: kate
    number_sections: true
    theme: spacelab
    toc: true
    toc_float: true
    code_folding: "hide"
  pdf_document:
    toc: true
subtitle: Cirugía digestiva, endocrina y trasplante de órganos abdominales (IMIB)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F)
```

```{r, warning=FALSE, message=FALSE}
source("Librerias.R")
source("utils.R")

covs <- readRDS("../RDSs/covs.RDS")
tse_lp <- readRDS("../RDSs/tse_lp.RDS")
tse_t1 <- readRDS("../RDSs/tse_t1.RDS")
tse_t3 <- readRDS("../RDSs/tse_t3.RDS")
tse_bilis <- readRDS("../RDSs/tse_bilis.RDS")
tse_stool <- readRDS("../RDSs/tse_stool.RDS")
tse_higado_sra <- readRDS("../RDSs/tse_higado_sra.RDS")
tse_bilis_sra <- readRDS("../RDSs/tse_bilis_sra.RDS")
tse_bilis_sra2 <- readRDS("../RDSs/tse_bilis_sra2.RDS")

```

# Cargar muestras

## Renombrar Phylum

```{r}
lp_nombres <- rownames(assay(agglomerateByRank(tse_lp, rank = "Phylum")))
t1_nombres <- rownames(assay(agglomerateByRank(tse_t1, rank = "Phylum")))
t3_nombres <- rownames(assay(agglomerateByRank(tse_t3, rank = "Phylum")))
sra_nombres_higado <- rownames(assay(agglomerateByRank(tse_higado_sra, rank = "Phylum")))
bilis_nombres <- rownames(assay(agglomerateByRank(tse_bilis, rank = "Phylum")))
stool_nombres <- rownames(assay(agglomerateByRank(tse_stool, rank = "Phylum")))
sra_nombres_bilis <- rownames(assay(agglomerateByRank(tse_bilis_sra, rank = "Phylum")))
sra_nombres_bilis2 <- rownames(assay(agglomerateByRank(tse_bilis_sra2, rank = "Phylum")))


setdiff(c(lp_nombres, t1_nombres, t3_nombres, bilis_nombres), c(sra_nombres_higado, stool_nombres, sra_nombres_bilis, sra_nombres_bilis2))
cat("\n")
setdiff(c(sra_nombres_higado, stool_nombres, sra_nombres_bilis, sra_nombres_bilis2), c(lp_nombres, t1_nombres, t3_nombres, bilis_nombres))

correspondencia_nombres <- c("Acidobacteriota" = "Acidobacteria",
                             "Actinobacteriota" = "Actinobacteria",
                             "Bacteroidota" = "Bacteroidetes",
                             "Gemmatimonadota" = "Gemmatimonadetes",
                             "Nitrospirota" = "Nitrospirae"
                             )

sra_nombres_higado <- ifelse(sra_nombres_higado %in% names(correspondencia_nombres), 
                      correspondencia_nombres[sra_nombres_higado], 
                      sra_nombres_higado)
stool_nombres <- ifelse(stool_nombres %in% names(correspondencia_nombres), 
                      correspondencia_nombres[stool_nombres], 
                      stool_nombres)
sra_nombres_bilis <- ifelse(sra_nombres_bilis %in% names(correspondencia_nombres), 
                      correspondencia_nombres[sra_nombres_bilis], 
                      sra_nombres_bilis)
sra_nombres_bilis2 <- ifelse(sra_nombres_bilis2 %in% names(correspondencia_nombres), 
                      correspondencia_nombres[sra_nombres_bilis2], 
                      sra_nombres_bilis2)

cat("\n")
setdiff(c(lp_nombres, t1_nombres, t3_nombres, bilis_nombres), c(sra_nombres_higado, stool_nombres, sra_nombres_bilis, sra_nombres_bilis2))
cat("\n")
setdiff(c(sra_nombres_higado, stool_nombres, sra_nombres_bilis, sra_nombres_bilis2), c(lp_nombres, t1_nombres, t3_nombres, bilis_nombres))

rowData(tse_higado_sra)$Phylum <- ifelse(rowData(tse_higado_sra)$Phylum %in% names(correspondencia_nombres),
                      correspondencia_nombres[rowData(tse_higado_sra)$Phylum],
                      rowData(tse_higado_sra)$Phylum)

rowData(tse_stool)$Phylum <- ifelse(rowData(tse_stool)$Phylum %in% names(correspondencia_nombres),
                      correspondencia_nombres[rowData(tse_stool)$Phylum],
                      rowData(tse_stool)$Phylum)

rowData(tse_bilis_sra)$Phylum <- ifelse(rowData(tse_bilis_sra)$Phylum %in% names(correspondencia_nombres),
                      correspondencia_nombres[rowData(tse_bilis_sra)$Phylum],
                      rowData(tse_bilis_sra)$Phylum)

rowData(tse_bilis_sra2)$Phylum <- ifelse(rowData(tse_bilis_sra2)$Phylum %in% names(correspondencia_nombres),
                      correspondencia_nombres[rowData(tse_bilis_sra2)$Phylum],
                      rowData(tse_bilis_sra2)$Phylum)

```

## Renombrar Class

```{r}
lp_nombres <- rownames(assay(agglomerateByRank(tse_lp, rank = "Class")))
t1_nombres <- rownames(assay(agglomerateByRank(tse_t1, rank = "Class")))
t3_nombres <- rownames(assay(agglomerateByRank(tse_t3, rank = "Class")))
sra_nombres_higado <- rownames(assay(agglomerateByRank(tse_higado_sra, rank = "Class")))
bilis_nombres <- rownames(assay(agglomerateByRank(tse_bilis, rank = "Class")))
stool_nombres <- rownames(assay(agglomerateByRank(tse_stool, rank = "Class")))
sra_nombres_bilis <- rownames(assay(agglomerateByRank(tse_bilis_sra, rank = "Class")))
sra_nombres_bilis2 <- rownames(assay(agglomerateByRank(tse_bilis_sra2, rank = "Class")))


setdiff(c(lp_nombres, t1_nombres, t3_nombres, bilis_nombres), c(sra_nombres_higado, stool_nombres, sra_nombres_bilis, sra_nombres_bilis2))
cat("\n")
setdiff(c(sra_nombres_higado, stool_nombres, sra_nombres_bilis, sra_nombres_bilis2), c(lp_nombres, t1_nombres, t3_nombres, bilis_nombres))

correspondencia_nombres <- c("Acidobacteriota" = "Acidobacteria",
                             "Actinobacteriota" = "Actinobacteria",
                             "Bacteroidota" = "Bacteroidetes",
                             "Gemmatimonadota" = "Gemmatimonadetes",
                             "Nitrospirota" = "Nitrospirae"
                             )

sra_nombres_higado <- ifelse(sra_nombres_higado %in% names(correspondencia_nombres), 
                      correspondencia_nombres[sra_nombres_higado], 
                      sra_nombres_higado)
stool_nombres <- ifelse(stool_nombres %in% names(correspondencia_nombres), 
                      correspondencia_nombres[stool_nombres], 
                      stool_nombres)
sra_nombres_bilis <- ifelse(sra_nombres_bilis %in% names(correspondencia_nombres), 
                      correspondencia_nombres[sra_nombres_bilis], 
                      sra_nombres_bilis)
sra_nombres_bilis2 <- ifelse(sra_nombres_bilis2 %in% names(correspondencia_nombres), 
                      correspondencia_nombres[sra_nombres_bilis2], 
                      sra_nombres_bilis2)

cat("\n")
setdiff(c(lp_nombres, t1_nombres, t3_nombres, bilis_nombres), c(sra_nombres_higado, stool_nombres, sra_nombres_bilis, sra_nombres_bilis2))
cat("\n")
setdiff(c(sra_nombres_higado, stool_nombres, sra_nombres_bilis, sra_nombres_bilis2), c(lp_nombres, t1_nombres, t3_nombres, bilis_nombres))

rowData(tse_higado_sra)$Class <- ifelse(rowData(tse_higado_sra)$Class %in% names(correspondencia_nombres),
                      correspondencia_nombres[rowData(tse_higado_sra)$Class],
                      rowData(tse_higado_sra)$Class)

rowData(tse_stool)$Class <- ifelse(rowData(tse_stool)$Class %in% names(correspondencia_nombres),
                      correspondencia_nombres[rowData(tse_stool)$Class],
                      rowData(tse_stool)$Class)

rowData(tse_bilis_sra)$Class <- ifelse(rowData(tse_bilis_sra)$Class %in% names(correspondencia_nombres),
                      correspondencia_nombres[rowData(tse_bilis_sra)$Class],
                      rowData(tse_bilis_sra)$Class)

rowData(tse_bilis_sra2)$Class <- ifelse(rowData(tse_bilis_sra2)$Class %in% names(correspondencia_nombres),
                      correspondencia_nombres[rowData(tse_bilis_sra2)$Class],
                      rowData(tse_bilis_sra2)$Class)

```

"‘standardize’ Standardize(or ‘z-score’) transformation scales data to zero mean and unit variance. This is used to bring features (or samples) to more comparable levels in terms of mean and scale of the values. This can enhance visualization and interpretation of the data"

```{r}
tse_lp <- transformAssay(tse_lp, assay.type = "counts", method = "relabundance")
tse_lp <- transformAssay(tse_lp, assay.type = "counts", method = "standardize")
tse_lp <- transformAssay(tse_lp, assay.type = "relabundance", method = "clr", pseudocount = TRUE)

tse_t1 <- transformAssay(tse_t1, assay.type = "counts", method = "relabundance")
tse_t1 <- transformAssay(tse_t1, assay.type = "counts", method = "standardize")
tse_t1 <- transformAssay(tse_t1, assay.type = "relabundance", method = "clr", pseudocount = TRUE)

tse_t3 <- transformAssay(tse_t3, assay.type = "counts", method = "relabundance")
tse_t3 <- transformAssay(tse_t3, assay.type = "counts", method = "standardize")
tse_t3 <- transformAssay(tse_t3, assay.type = "relabundance", method = "clr", pseudocount = TRUE)

tse_higado_sra <- transformAssay(tse_higado_sra, assay.type = "counts", method = "relabundance")
tse_higado_sra <- transformAssay(tse_higado_sra, assay.type = "counts", method = "standardize")
tse_higado_sra <- transformAssay(tse_higado_sra, assay.type = "relabundance", method = "clr", pseudocount = TRUE)

tse_bilis <- transformAssay(tse_bilis, assay.type = "counts", method = "relabundance")
tse_bilis <- transformAssay(tse_bilis, assay.type = "counts", method = "standardize")
tse_bilis <- transformAssay(tse_bilis, assay.type = "relabundance", method = "clr", pseudocount = TRUE)

tse_bilis_sra <- transformAssay(tse_bilis_sra, assay.type = "counts", method = "relabundance")
tse_bilis_sra <- transformAssay(tse_bilis_sra, assay.type = "counts", method = "standardize")
tse_bilis_sra <- transformAssay(tse_bilis_sra, assay.type = "relabundance", method = "clr", pseudocount = TRUE)

tse_bilis_sra2 <- transformAssay(tse_bilis_sra2, assay.type = "counts", method = "relabundance")
tse_bilis_sra2 <- transformAssay(tse_bilis_sra2, assay.type = "counts", method = "standardize")
tse_bilis_sra2 <- transformAssay(tse_bilis_sra2, assay.type = "relabundance", method = "clr", pseudocount = TRUE)

```

# Community composition

## Relative abundance

```{r, fig.width=8, fig.height= 5}
abundancias <- abundancias_relativas_totales(list(tse_lp = tse_lp, 
                                                  tse_bilis = tse_bilis,
                                                  tse_t1 = tse_t1,
                                                  tse_t3 = tse_t3,
                                                  tse_higado_sra= tse_higado_sra,
                                                  tse_stool = tse_stool,
                                                  tse_bilis_sra = tse_bilis_sra,
                                                  tse_bilis_sra2 = tse_bilis_sra2),
                                             ranking = "Phylum",
                                             top = 5)


abundancias <- abundancias %>%
    group_by(Taxon) %>%
    mutate(order_TSE_lp = ifelse(TSE == "tse_lp", Relative_Abundance, NA)) %>%
    arrange(desc(order_TSE_lp)) %>%
    ungroup() %>%
    mutate(taxa_sub = factor(taxa_sub, levels = c(unique(taxa_sub[taxa_sub != "Other"]), "Other")),
           TSE = factor(TSE, levels = c("tse_lp", "tse_t1", "tse_t3", "tse_bilis", "tse_bilis_sra", "tse_bilis_sra2", "tse_higado_sra",  "tse_stool")))

plotear_abundancia_relativa(abundancias, "Phylum")
```

```{r}
tse_lp_phylum <- agglomerateByRank(tse_lp, "Phylum")

counts_lp <- as.data.frame(t(assay(tse_lp_phylum, "counts"))) %>% 
    rownames_to_column(var = "sample")




covariates_lp <- colData(tse_lp_phylum) %>%
    as.data.frame()%>%
    rownames_to_column(var = "sample")

dplyr::left_join(counts_lp, covariates_lp, by = "sample") %>%
    dplyr::select(intersect(names(.), names(counts_lp)), Supervivencia, -sample) %>%
    pivot_longer(-Supervivencia, names_to = "Taxon", values_to = "Abundance") %>%
    group_by(Supervivencia, Taxon) %>%
    summarize(suma = sum(Abundance)) %>%
    group_by(Supervivencia) %>%
    mutate(suma_condition = sum(suma),
           relative_abundance = suma / suma_condition) %>%
    ggplot(aes(x=Supervivencia, y=relative_abundance, fill = Taxon)) +
    geom_bar(stat = "identity") +
    #scale_fill_paletteer_d("ggthemes::Tableau_10") + 
    labs(title = "Phylum Relative Abundance", x = "Taxons", y = "Relative Abundance") +
    theme_minimal() +
    theme(
        # axis.text.x = element_blank(),
        # axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        legend.title = element_blank(),
        legend.position = "none",  # Posiciona la leyenda abajo
        #legend.box.margin = margin(t = 10),  # Añade espacio superior para la leyenda
        legend.spacing.x = unit(0.5, 'cm'),  # Ajusta espacio entre elementos de la leyenda
        plot.title = element_text(hjust = 0.5, vjust = 10, size = 16),  # Sube el título y ajusta el tamaño
       #plot.margin = margin(t = 20)  # Añade margen superior para el título
        ) 


```

```{r, fig.width=10, fig.height=10}
plotear_diferencias_relabundance(tse_lp, "Phylum", "AR")
```

```{r, fig.width=10, fig.height=10}
plotear_diferencias_relabundance(tse_lp, "Class", "Donación")
```

```{r, fig.width=10, fig.height=5}
abundancias <- abundancias_relativas_totales(list(tse_lp = tse_lp, 
                                                  tse_bilis = tse_bilis,
                                                  tse_t1 = tse_t1,
                                                  tse_t3 = tse_t3,
                                                  tse_higado_sra= tse_higado_sra,
                                                  tse_stool = tse_stool,
                                                  tse_bilis_sra = tse_bilis_sra,
                                                  tse_bilis_sra2 = tse_bilis_sra2),
                                             ranking = "Class",
                                             top = 5)


abundancias <- abundancias %>%
    group_by(Taxon) %>%
    mutate(order_TSE_lp = ifelse(TSE == "tse_lp", Relative_Abundance, NA)) %>%
    arrange(desc(order_TSE_lp)) %>%
    ungroup() %>%
    mutate(taxa_sub = factor(taxa_sub, levels = c(unique(taxa_sub[taxa_sub != "Other"]), "Other")),
           TSE = factor(TSE, levels = c("tse_lp", "tse_t1", "tse_t3", "tse_bilis", "tse_bilis_sra", "tse_bilis_sra2", "tse_higado_sra", "tse_stool")))


plotear_abundancia_relativa(abundancias, "Class")
```

```{r, fig.width=10, fig.height=5}
abundancias <- abundancias_relativas_totales(list(tse_lp = tse_lp, 
                                                  tse_bilis = tse_bilis,
                                                  tse_t1 = tse_t1,
                                                  tse_t3 = tse_t3,
                                                  tse_higado_sra= tse_higado_sra,
                                                  tse_stool = tse_stool,
                                                  tse_bilis_sra = tse_bilis_sra),
                                             ranking = "Order",
                                             top = 5)


abundancias <- abundancias %>%
    group_by(Taxon) %>%
    mutate(order_TSE_lp = ifelse(TSE == "tse_lp", Relative_Abundance, NA)) %>%
    arrange(desc(order_TSE_lp)) %>%
    ungroup() %>%
    mutate(taxa_sub = factor(taxa_sub, levels = c(unique(taxa_sub[taxa_sub != "Other"]), "Other")),
           TSE = factor(TSE, levels = c("tse_lp", "tse_t1", "tse_t3", "tse_bilis", "tse_bilis_sra", "tse_higado_sra", "tse_stool")))


plotear_abundancia_relativa(abundancias, "Order")
```

```{r, fig.width=10, fig.height=5}
abundancias <- abundancias_relativas_totales(list(tse_lp = tse_lp, 
                                                  tse_bilis = tse_bilis,
                                                  tse_t1 = tse_t1,
                                                  tse_t3 = tse_t3,
                                                  tse_higado_sra= tse_higado_sra,
                                                  tse_stool = tse_stool,
                                                  tse_bilis_sra = tse_bilis_sra),
                                             ranking = "Family",
                                             top = 5)


abundancias <- abundancias %>%
    group_by(Taxon) %>%
    mutate(order_TSE_lp = ifelse(TSE == "tse_lp", Relative_Abundance, NA)) %>%
    arrange(desc(order_TSE_lp)) %>%
    ungroup() %>%
    mutate(taxa_sub = factor(taxa_sub, levels = c(unique(taxa_sub[taxa_sub != "Other"]), "Other")),
           TSE = factor(TSE, levels = c("tse_lp", "tse_t1", "tse_t3", "tse_bilis", "tse_bilis_sra", "tse_higado_sra", "tse_stool")))


plotear_abundancia_relativa(abundancias, "Family")
```

```{r, fig.width=10, fig.height=5}
abundancias <- abundancias_relativas_totales(list(tse_lp = tse_lp, 
                                                  tse_bilis = tse_bilis,
                                                  tse_t1 = tse_t1,
                                                  tse_t3 = tse_t3,
                                                  tse_higado_sra= tse_higado_sra,
                                                  tse_stool = tse_stool,
                                                  tse_bilis_sra = tse_bilis_sra),
                                             ranking = "Genus",
                                             top = 5)


abundancias <- abundancias %>%
    group_by(Taxon) %>%
    mutate(order_TSE_lp = ifelse(TSE == "tse_lp", Relative_Abundance, NA)) %>%
    arrange(desc(order_TSE_lp)) %>%
    ungroup() %>%
    mutate(taxa_sub = factor(taxa_sub, levels = c(unique(taxa_sub[taxa_sub != "Other"]), "Other")),
           TSE = factor(TSE, levels = c("tse_lp", "tse_t1", "tse_t3", "tse_bilis", "tse_bilis_sra", "tse_higado_sra", "tse_stool")))


plotear_abundancia_relativa(abundancias, "Genus")
```

```{r}
abundancias <- abundancias_relativas_totales(list(tse_lp = tse_lp, 
                                                  tse_bilis = tse_bilis,
                                                  tse_t1 = tse_t1,
                                                  #tse_higado_sra= tse_higado_sra,
                                                  tse_stool = tse_stool
                                                  ),
                                             ranking = "Species",
                                             top = 5)


abundancias <- abundancias %>%
    group_by(Taxon) %>%
    mutate(order_TSE_lp = ifelse(TSE == "tse_lp", Relative_Abundance, NA)) %>%
    arrange(desc(order_TSE_lp)) %>%
    ungroup() %>%
    mutate(taxa_sub = factor(taxa_sub, levels = c(unique(taxa_sub[taxa_sub != "Other"]), "Other")),
           TSE = factor(TSE, levels = c("tse_lp", "tse_t1", "tse_bilis", "tse_higado_sra", "tse_stool")))


plotear_abundancia_relativa(abundancias, "Species")
```

## Differential Abundance Analysis (DAA)

<https://bioconductor.org/packages/devel/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#differential-expression-analysis>

### DESEq2

#### Species

```{r, warning=FALSE, message=FALSE}
taxa_relevante_species <- data.frame()

# Agglomerate by genus and subset by prevalence
tse_prevalente <- subsetByPrevalent(tse_lp, rank = "Species", prevalence = 0.1)
tse_rara <- subsetByRare(tse_lp, rank = "Species", prevalence = 0.23)

# Transform count assay to relative abundances
tse_prevalente <- transformAssay(tse_prevalente, assay.type = "counts", method = "relabundance")
tse_rara <- transformAssay(tse_rara, assay.type = "counts", method = "relabundance")

dda_deseq_prevalente <- deseq_flr(tse_prevalente, rara = T, prevalence = "prevalente", taxa_relevante = taxa_relevante_species)
dda_deseq_rara <- deseq_flr(tse_rara, rara = T, prevalence = "rara", taxa_relevante = taxa_relevante_species)
```


```{r, fig.width=10, fig.height=7, warning=FALSE}
do.call(grid.arrange, c(volcanos_deseq_flr(dda_deseq_prevalente$objetos_deseq2), ncol= 2))
do.call(grid.arrange, c(volcanos_deseq_flr(dda_deseq_rara$objetos_deseq2), ncol= 2))
```

<https://cran.r-project.org/web/packages/forestploter/vignettes/forestploter-intro.html>

```{r, fig.width=10, fig.height=10, warning=FALSE}
# 
# df_filtrado <- df_filtrado %>%
#     dplyr::arrange(desc(abs(log2FoldChange)), padj) %>% head(20)
# 
# relabundance <- as.data.frame(t(assay(tse_rara, "relabundance"))) %>% 
#     rownames_to_column(var = "sample") %>%
#     dplyr::select(all_of(df_filtrado$code), sample)
# 
# covariates <- colData(tse_rara) %>%
#     as.data.frame() %>%
#     rownames_to_column(var = "sample")
# 
# p <- dplyr::left_join(relabundance, covariates, by = "sample") %>%
#     dplyr::select(intersect(names(.), names(relabundance)), all_of("Biliary_complications"), -sample) %>%
#     pivot_longer(-all_of("Biliary_complications"), names_to = "Taxon", values_to = "Abundance") %>%
#     ggplot(aes_string(x="Biliary_complications", y="Abundance", fill = "Taxon")) +
#     geom_boxplot() +
#     geom_jitter(aes_string(x="Biliary_complications", y="Abundance"), size = 0.5) +
#     labs(title = paste("Relative Abundance at", "taxa", "level"),
#          x = "Biliary_complications", 
#          y = "Relative Abundance") +
#     theme(
#         axis.title.y = element_blank(),
#         legend.title = element_blank(),
#         legend.position = "none"
#     ) +
#     stat_compare_means() +
#     scale_y_log10() +
#     facet_wrap(~ reorder(Taxon, -Abundance))
# 
# p
```

```{r, fig.width=8, fig.height=4}
lista <- list()
for (i in dda_deseq_prevalente$taxa_relevante$variable) {
    
    microorganismos <- dda_deseq_prevalente$taxa_relevante %>%
        dplyr::filter(variable == i) %>%
        pull(code)
    
    lista[[i]] <- microorganismos
    
}

upset(fromList(lista), 
      order.by = "freq",
      sets.x.label = "Taxonomy by variable",
      mainbar.y.label = "Taxonomy",
      point.size = 3.5, 
      line.size = 1.5,
      text.scale = c(1.3, 1.3, 1, 1, 1.5, 1.3),

)
```

```{r, fig.width=8, fig.height=4}
lista <- list()
for (i in dda_deseq_rara$taxa_relevante$variable) {
    
    microorganismos <- dda_deseq_rara$taxa_relevante %>%
        dplyr::filter(variable == i) %>%
        pull(code)
    
    lista[[i]] <- microorganismos
    
}

upset(fromList(lista), 
      order.by = "freq",
      sets.x.label = "Taxonomy by variable",
      mainbar.y.label = "Taxonomy",
      point.size = 3.5, 
      line.size = 1.5,
      text.scale = c(1.3, 1.3, 1, 1, 1.5, 1.3),

)
```

#### Genus

```{r, warning=FALSE, message=FALSE}
taxa_relevante_genus <- data.frame()

# Agglomerate by genus and subset by prevalence
tse_prevalente <- subsetByPrevalent(tse_lp, rank = "Genus", prevalence = 0.1)
tse_rara <- subsetByRare(tse_lp, rank = "Genus", prevalence = 0.23)

# Transform count assay to relative abundances
tse_prevalente <- transformAssay(tse_prevalente, assay.type = "counts", method = "relabundance")
tse_rara <- transformAssay(tse_rara, assay.type = "counts", method = "relabundance")

dda_deseq_prevalente <- deseq_flr(tse_prevalente, rara = T, prevalence = "prevalente", taxa_relevante = taxa_relevante_genus)
dda_deseq_rara <- deseq_flr(tse_rara, rara = T, prevalence = "rara", taxa_relevante = taxa_relevante_genus)
```


```{r, fig.width=10, fig.height=7, warning=FALSE}
do.call(grid.arrange, c(volcanos_deseq_flr(dda_deseq_prevalente$objetos_deseq2), ncol= 2))
do.call(grid.arrange, c(volcanos_deseq_flr(dda_deseq_rara$objetos_deseq2), ncol= 2))
```

<https://cran.r-project.org/web/packages/forestploter/vignettes/forestploter-intro.html>

```{r, fig.width=10, fig.height=10, warning=FALSE}
# 
# df_filtrado <- df_filtrado %>%
#     dplyr::arrange(desc(abs(log2FoldChange)), padj) %>% head(20)
# 
# relabundance <- as.data.frame(t(assay(tse_rara, "relabundance"))) %>% 
#     rownames_to_column(var = "sample") %>%
#     dplyr::select(all_of(df_filtrado$code), sample)
# 
# covariates <- colData(tse_rara) %>%
#     as.data.frame() %>%
#     rownames_to_column(var = "sample")
# 
# p <- dplyr::left_join(relabundance, covariates, by = "sample") %>%
#     dplyr::select(intersect(names(.), names(relabundance)), all_of("Biliary_complications"), -sample) %>%
#     pivot_longer(-all_of("Biliary_complications"), names_to = "Taxon", values_to = "Abundance") %>%
#     ggplot(aes_string(x="Biliary_complications", y="Abundance", fill = "Taxon")) +
#     geom_boxplot() +
#     geom_jitter(aes_string(x="Biliary_complications", y="Abundance"), size = 0.5) +
#     labs(title = paste("Relative Abundance at", "taxa", "level"),
#          x = "Biliary_complications", 
#          y = "Relative Abundance") +
#     theme(
#         axis.title.y = element_blank(),
#         legend.title = element_blank(),
#         legend.position = "none"
#     ) +
#     stat_compare_means() +
#     scale_y_log10() +
#     facet_wrap(~ reorder(Taxon, -Abundance))
# 
# p
```

```{r, fig.width=8, fig.height=4}
lista <- list()
for (i in dda_deseq_prevalente$taxa_relevante$variable) {
    
    microorganismos <- dda_deseq_prevalente$taxa_relevante %>%
        dplyr::filter(variable == i) %>%
        pull(code)
    
    lista[[i]] <- microorganismos
    
}

upset(fromList(lista), 
      order.by = "freq",
      sets.x.label = "Taxonomy by variable",
      mainbar.y.label = "Taxonomy",
      point.size = 3.5, 
      line.size = 1.5,
      text.scale = c(1.3, 1.3, 1, 1, 1.5, 1.3),

)
```

```{r}
dda_deseq_prevalente$taxa_relevante %>%
  group_by(code) %>%                        # Agrupar por la columna "code"
  filter(all(variable == "AR")) %>% 
  ungroup()


```


```{r, fig.width=8, fig.height=4}
lista <- list()
for (i in dda_deseq_rara$taxa_relevante$variable) {
    
    microorganismos <- dda_deseq_rara$taxa_relevante %>%
        dplyr::filter(variable == i) %>%
        pull(code)
    
    lista[[i]] <- microorganismos
    
}

upset(fromList(lista), 
      order.by = "freq",
      sets.x.label = "Taxonomy by variable",
      mainbar.y.label = "Taxonomy",
      point.size = 3.5, 
      line.size = 1.5,
      text.scale = c(1.3, 1.3, 1, 1, 1.5, 1.3),

)
```

```{r}
Reduce(intersect, lista[1:3])

rowData(tse_lp) %>%
    as.data.frame() %>%
    filter(Species == "propinquum")

```

### MaAsLin2

<https://huttenhower.sph.harvard.edu/maaslin/>

#### Species

```{r, warning=FALSE, message=FALSE}
# Agglomerate by genus and subset by prevalence
tse_prevalente <- subsetByPrevalent(tse_lp, rank = "Species", prevalence = 0.1)
tse_rara <- subsetByRare(tse_lp, rank = "Species", prevalence = 0.23)

# Transform count assay to relative abundances
tse_prevalente <- transformAssay(tse_prevalente, assay.type = "counts", method = "relabundance")
tse_rara <- transformAssay(tse_rara, assay.type = "counts", method = "relabundance")
```

```{r, eval=FALSE}

variables_salida <- c("Supervivencia", "AR", "AHT", "Biliary_complications")

for (i in variables_salida) {
    
    output_path <- paste0("../maaslin2/species/prevalente/", i, "/")
    dir.create(output_path, recursive = TRUE, showWarnings = FALSE)
    
    fit_data2 <- Maaslin2(
        input_data = as.data.frame(t(assay(tse_prevalente))), 
        input_metadata = as.data.frame(colData(tse_prevalente)), 
        output = output_path,
        fixed_effects = c("Batch", i), 
        reference = c(paste0("Batch,", "RUN 3"), paste0(i, ",", ifelse(i == "Supervivencia", "SI", "NO"))), 
        cores = 4)
    
}

for (i in variables_salida) {
    
    output_path <- paste0("../maaslin2/species/rara/", i, "/")
    dir.create(output_path, recursive = TRUE, showWarnings = FALSE)
    
    fit_data2 <- Maaslin2(
        input_data = as.data.frame(t(assay(tse_rara))), 
        input_metadata = as.data.frame(colData(tse_rara)), 
        output = output_path,
        fixed_effects = c("Batch", i), 
        reference = c(paste0("Batch,", "RUN 3"), paste0(i, ",", ifelse(i == "Supervivencia", "SI", "NO"))), 
        cores = 4)
    
}


```

```{r}
dda_maaslin_prevalente <- list()

variables_salida <- c("Supervivencia", "AR", "AHT", "Biliary_complications")

for (i in variables_salida) {
    
    maslin_objeto <- read_delim(paste0("../maaslin2/species/prevalente/", i, "/all_results.tsv")) %>% 
    dplyr::filter(value != "RUN 1" , value != "RUN 2") 
    
    dda_maaslin_prevalente[[i]] <- maslin_objeto
    
}

dda_maaslin_rara<- list()

for (i in variables_salida) {
    
    maslin_objeto <- read_delim(paste0("../maaslin2/species/rara/", i, "/all_results.tsv")) %>% 
    dplyr::filter(value != "RUN 1" , value != "RUN 2") 
    
    dda_maaslin_rara[[i]] <- maslin_objeto
    
}
```


```{r, fig.width=10, fig.height=7, warning=FALSE}
do.call(grid.arrange, c(volcanos_maaslin_flr(dda_maaslin_prevalente), ncol= 2))
do.call(grid.arrange, c(volcanos_maaslin_flr(dda_maaslin_rara), ncol= 2))
```

<https://cran.r-project.org/web/packages/forestploter/vignettes/forestploter-intro.html>

```{r, fig.width=10, fig.height=10, warning=FALSE}
# 
# df_filtrado <- df_filtrado %>%
#     dplyr::arrange(desc(abs(log2FoldChange)), padj) %>% head(20)
# 
# relabundance <- as.data.frame(t(assay(tse_rara, "relabundance"))) %>% 
#     rownames_to_column(var = "sample") %>%
#     dplyr::select(all_of(df_filtrado$code), sample)
# 
# covariates <- colData(tse_rara) %>%
#     as.data.frame() %>%
#     rownames_to_column(var = "sample")
# 
# p <- dplyr::left_join(relabundance, covariates, by = "sample") %>%
#     dplyr::select(intersect(names(.), names(relabundance)), all_of("Biliary_complications"), -sample) %>%
#     pivot_longer(-all_of("Biliary_complications"), names_to = "Taxon", values_to = "Abundance") %>%
#     ggplot(aes_string(x="Biliary_complications", y="Abundance", fill = "Taxon")) +
#     geom_boxplot() +
#     geom_jitter(aes_string(x="Biliary_complications", y="Abundance"), size = 0.5) +
#     labs(title = paste("Relative Abundance at", "taxa", "level"),
#          x = "Biliary_complications", 
#          y = "Relative Abundance") +
#     theme(
#         axis.title.y = element_blank(),
#         legend.title = element_blank(),
#         legend.position = "none"
#     ) +
#     stat_compare_means() +
#     scale_y_log10() +
#     facet_wrap(~ reorder(Taxon, -Abundance))
# 
# p
```

```{r, fig.width=8, fig.height=4}
taxa_relevante_maaslin_especie <- data.frame()

for (i in names(dda_maaslin_prevalente)) {
        
    i_filtrado <- dda_maaslin_prevalente[[i]] %>%
        dplyr::filter(pval < 0.05) %>%
        dplyr::select(feature, coef, stderr, pval)
    
    i_filtrado$variable <- i
    
    i_filtrado$prevalencia <- "prevalente"
    
    for (j in 1:nrow(i_filtrado)) {
        
        if (i_filtrado[j, "coef"] > 0) {
            
            i_filtrado[j, "biosis"] <- "Hiperbiosis"
            
        } else {
            
            i_filtrado[j, "biosis"] <- "Hipobiosis"
            
        }
        
    }
    
    if (nrow(taxa_relevante_maaslin_especie) == 0){
            
        taxa_relevante_maaslin_especie <- i_filtrado
        
    } else {
        
        taxa_relevante_maaslin_especie <- rbind(taxa_relevante_maaslin_especie, i_filtrado)
        
    }
    
}

lista <- list()
for (i in taxa_relevante_maaslin_especie$variable) {
    
    microorganismos <- taxa_relevante_maaslin_especie %>%
        dplyr::filter(variable == i) %>%
        pull(feature)
    
    lista[[i]] <- microorganismos
    
}

upset(fromList(lista), 
      order.by = "freq",
      sets.x.label = "Taxonomy by variable",
      mainbar.y.label = "Taxonomy",
      point.size = 3.5, 
      line.size = 1.5,
      text.scale = c(1.3, 1.3, 1, 1, 1.5, 1.3),

)

```

```{r, fig.width=8, fig.height=4}
taxa_relevante_maaslin_especie <- data.frame()

for (i in names(dda_maaslin_rara)) {
        
    i_filtrado <- dda_maaslin_rara[[i]] %>%
        dplyr::filter(pval < 0.05) %>%
        dplyr::select(feature, coef, stderr, pval)
    
    i_filtrado$variable <- i
    
    i_filtrado$prevalencia <- "prevalente"
    
    for (j in 1:nrow(i_filtrado)) {
        
        if (i_filtrado[j, "coef"] > 0) {
            
            i_filtrado[j, "biosis"] <- "Hiperbiosis"
            
        } else {
            
            i_filtrado[j, "biosis"] <- "Hipobiosis"
            
        }
        
    }
    
    if (nrow(taxa_relevante_maaslin_especie) == 0){
            
        taxa_relevante_maaslin_especie <- i_filtrado
        
    } else {
        
        taxa_relevante_maaslin_especie <- rbind(taxa_relevante_maaslin_especie, i_filtrado)
        
    }
    
}

lista <- list()
for (i in taxa_relevante_maaslin_especie$variable) {
    
    microorganismos <- taxa_relevante_maaslin_especie %>%
        dplyr::filter(variable == i) %>%
        pull(feature)
    
    lista[[i]] <- microorganismos
    
}

upset(fromList(lista), 
      order.by = "freq",
      sets.x.label = "Taxonomy by variable",
      mainbar.y.label = "Taxonomy",
      point.size = 3.5, 
      line.size = 1.5,
      text.scale = c(1.3, 1.3, 1, 1, 1.5, 1.3),

)

```

#### Genus

```{r, warning=FALSE, message=FALSE}
# Agglomerate by genus and subset by prevalence
tse_prevalente <- subsetByPrevalent(tse_lp, rank = "Genus", prevalence = 0.1)
tse_rara <- subsetByRare(tse_lp, rank = "Genus", prevalence = 0.23)

# Transform count assay to relative abundances
tse_prevalente <- transformAssay(tse_prevalente, assay.type = "counts", method = "relabundance")
tse_rara <- transformAssay(tse_rara, assay.type = "counts", method = "relabundance")
```

```{r, eval=FALSE}

variables_salida <- c("Supervivencia", "AR", "AHT", "Biliary_complications")

for (i in variables_salida) {
    
    output_path <- paste0("../maaslin2/genus/prevalente/", i, "/")
    dir.create(output_path, recursive = TRUE, showWarnings = FALSE)
    
    fit_data2 <- Maaslin2(
        input_data = as.data.frame(t(assay(tse_prevalente))), 
        input_metadata = as.data.frame(colData(tse_prevalente)), 
        output = output_path,
        fixed_effects = c("Batch", i), 
        reference = c(paste0("Batch,", "RUN 3"), paste0(i, ",", ifelse(i == "Supervivencia", "SI", "NO"))), 
        cores = 4)
    
}

for (i in variables_salida) {
    
    output_path <- paste0("../maaslin2/genus/rara/", i, "/")
    dir.create(output_path, recursive = TRUE, showWarnings = FALSE)
    
    fit_data2 <- Maaslin2(
        input_data = as.data.frame(t(assay(tse_rara))), 
        input_metadata = as.data.frame(colData(tse_rara)), 
        output = output_path,
        fixed_effects = c("Batch", i), 
        reference = c(paste0("Batch,", "RUN 3"), paste0(i, ",", ifelse(i == "Supervivencia", "SI", "NO"))), 
        cores = 4)
    
}


```

```{r}
dda_maaslin_prevalente <- list()

variables_salida <- c("Supervivencia", "AR", "AHT", "Biliary_complications")

for (i in variables_salida) {
    
    maslin_objeto <- read_delim(paste0("../maaslin2/genus/prevalente/", i, "/all_results.tsv")) %>% 
    dplyr::filter(value != "RUN 1" , value != "RUN 2") 
    
    dda_maaslin_prevalente[[i]] <- maslin_objeto
    
}

dda_maaslin_rara <- list()

for (i in variables_salida) {
    
    maslin_objeto <- read_delim(paste0("../maaslin2/genus/rara/", i, "/all_results.tsv")) %>% 
    dplyr::filter(value != "RUN 1" , value != "RUN 2") 
    
    dda_maaslin_rara[[i]] <- maslin_objeto
    
}
```


```{r, fig.width=10, fig.height=7, warning=FALSE}
do.call(grid.arrange, c(volcanos_maaslin_flr(dda_maaslin_prevalente), ncol= 2))
do.call(grid.arrange, c(volcanos_maaslin_flr(dda_maaslin_rara), ncol= 2))
```

<https://cran.r-project.org/web/packages/forestploter/vignettes/forestploter-intro.html>

```{r, fig.width=10, fig.height=10, warning=FALSE}
# 
# df_filtrado <- df_filtrado %>%
#     dplyr::arrange(desc(abs(log2FoldChange)), padj) %>% head(20)
# 
# relabundance <- as.data.frame(t(assay(tse_rara, "relabundance"))) %>% 
#     rownames_to_column(var = "sample") %>%
#     dplyr::select(all_of(df_filtrado$code), sample)
# 
# covariates <- colData(tse_rara) %>%
#     as.data.frame() %>%
#     rownames_to_column(var = "sample")
# 
# p <- dplyr::left_join(relabundance, covariates, by = "sample") %>%
#     dplyr::select(intersect(names(.), names(relabundance)), all_of("Biliary_complications"), -sample) %>%
#     pivot_longer(-all_of("Biliary_complications"), names_to = "Taxon", values_to = "Abundance") %>%
#     ggplot(aes_string(x="Biliary_complications", y="Abundance", fill = "Taxon")) +
#     geom_boxplot() +
#     geom_jitter(aes_string(x="Biliary_complications", y="Abundance"), size = 0.5) +
#     labs(title = paste("Relative Abundance at", "taxa", "level"),
#          x = "Biliary_complications", 
#          y = "Relative Abundance") +
#     theme(
#         axis.title.y = element_blank(),
#         legend.title = element_blank(),
#         legend.position = "none"
#     ) +
#     stat_compare_means() +
#     scale_y_log10() +
#     facet_wrap(~ reorder(Taxon, -Abundance))
# 
# p
```

```{r, fig.width=8, fig.height=4}
taxa_relevante_maaslin_especie <- data.frame()

for (i in names(dda_maaslin_prevalente)) {
        
    i_filtrado <- dda_maaslin_prevalente[[i]] %>%
        dplyr::filter(pval < 0.05) %>%
        dplyr::select(feature, coef, stderr, pval)
    
    i_filtrado$variable <- i
    
    i_filtrado$prevalencia <- "prevalente"
    
    for (j in 1:nrow(i_filtrado)) {
        
        if (i_filtrado[j, "coef"] > 0) {
            
            i_filtrado[j, "biosis"] <- "Hiperbiosis"
            
        } else {
            
            i_filtrado[j, "biosis"] <- "Hipobiosis"
            
        }
        
    }
    
    if (nrow(taxa_relevante_maaslin_especie) == 0){
            
        taxa_relevante_maaslin_especie <- i_filtrado
        
    } else {
        
        taxa_relevante_maaslin_especie <- rbind(taxa_relevante_maaslin_especie, i_filtrado)
        
    }
    
}

lista <- list()
for (i in taxa_relevante_maaslin_especie$variable) {
    
    microorganismos <- taxa_relevante_maaslin_especie %>%
        dplyr::filter(variable == i) %>%
        pull(feature)
    
    lista[[i]] <- microorganismos
    
}

upset(fromList(lista), 
      order.by = "freq",
      sets.x.label = "Taxonomy by variable",
      mainbar.y.label = "Taxonomy",
      point.size = 3.5, 
      line.size = 1.5,
      text.scale = c(1.3, 1.3, 1, 1, 1.5, 1.3),

)

```

```{r, fig.width=8, fig.height=4}
taxa_relevante_maaslin_especie <- data.frame()

for (i in names(dda_maaslin_rara)) {
        
    i_filtrado <- dda_maaslin_rara[[i]] %>%
        dplyr::filter(pval < 0.05) %>%
        dplyr::select(feature, coef, stderr, pval)
    
    i_filtrado$variable <- i
    
    i_filtrado$prevalencia <- "prevalente"
    
    for (j in 1:nrow(i_filtrado)) {
        
        if (i_filtrado[j, "coef"] > 0) {
            
            i_filtrado[j, "biosis"] <- "Hiperbiosis"
            
        } else {
            
            i_filtrado[j, "biosis"] <- "Hipobiosis"
            
        }
        
    }
    
    if (nrow(taxa_relevante_maaslin_especie) == 0){
            
        taxa_relevante_maaslin_especie <- i_filtrado
        
    } else {
        
        taxa_relevante_maaslin_especie <- rbind(taxa_relevante_maaslin_especie, i_filtrado)
        
    }
    
}

lista <- list()
for (i in taxa_relevante_maaslin_especie$variable) {
    
    microorganismos <- taxa_relevante_maaslin_especie %>%
        dplyr::filter(variable == i) %>%
        pull(feature)
    
    lista[[i]] <- microorganismos
    
}

upset(fromList(lista), 
      order.by = "freq",
      sets.x.label = "Taxonomy by variable",
      mainbar.y.label = "Taxonomy",
      point.size = 3.5, 
      line.size = 1.5,
      text.scale = c(1.3, 1.3, 1, 1, 1.5, 1.3),

)

```

### ALDEX2

<https://bioconductor.org/packages/release/bioc/vignettes/ALDEx2/inst/doc/ALDEx2_vignette.html>

#### Species

```{r, warning=FALSE, message=FALSE}
# Agglomerate by genus and subset by prevalence
tse_prevalente <- subsetByPrevalent(tse_lp, rank = "Species", prevalence = 0.1)
tse_rara <- subsetByRare(tse_lp, rank = "Species", prevalence = 0.23)

# Transform count assay to relative abundances
tse_prevalente <- transformAssay(tse_prevalente, assay.type = "counts", method = "relabundance")
tse_rara <- transformAssay(tse_rara, assay.type = "counts", method = "relabundance")
```

```{r}
dda_aldex_prevalente <- aldex_flr(objeto_tse = tse_prevalente, prevalence = "prevalente", taxa_relevante = taxa_relevante_species)

dda_aldex_rara <- aldex_flr(objeto_tse = tse_rara, prevalence = "rara", taxa_relevante = taxa_relevante_species)
```

```{r, fig.width=10, fig.height=7, warning=FALSE}
do.call(grid.arrange, c(volcanos_aldex_flr(objetos_aldex = dda_aldex_prevalente$objetos_aldex), ncol= 2))
do.call(grid.arrange, c(volcanos_aldex_flr(objetos_aldex = dda_aldex_rara$objetos_aldex), ncol= 2))
```

<https://cran.r-project.org/web/packages/forestploter/vignettes/forestploter-intro.html>

```{r, fig.width=8, fig.height=4}

lista <- list()
for (i in dda_aldex_prevalente$taxa_relevante$variable) {
    
    microorganismos <- dda_aldex_prevalente$taxa_relevante %>%
        dplyr::filter(variable == i) %>%
        pull(code)
    
    lista[[i]] <- microorganismos
    
}

upset(fromList(lista), 
      order.by = "freq",
      sets.x.label = "Taxonomy by variable",
      mainbar.y.label = "Taxonomy",
      point.size = 3.5, 
      line.size = 1.5,
      text.scale = c(1.3, 1.3, 1, 1, 1.5, 1.3),

)

```

#### Genus

```{r, warning=FALSE, message=FALSE}
# Agglomerate by genus and subset by prevalence
tse_prevalente <- subsetByPrevalent(tse_lp, rank = "Genus", prevalence = 0.1)
tse_rara <- subsetByRare(tse_lp, rank = "Genus", prevalence = 0.23)

# Transform count assay to relative abundances
tse_prevalente <- transformAssay(tse_prevalente, assay.type = "counts", method = "relabundance")
tse_rara <- transformAssay(tse_rara, assay.type = "counts", method = "relabundance")
```

```{r}
dda_aldex_prevalente <- aldex_flr(objeto_tse = tse_prevalente, prevalence = "prevalente", taxa_relevante = taxa_relevante_species)

dda_aldex_rara <- aldex_flr(objeto_tse = tse_rara, prevalence = "rara", taxa_relevante = taxa_relevante_species)
```

```{r, fig.width=10, fig.height=7, warning=FALSE}
do.call(grid.arrange, c(volcanos_aldex_flr(objetos_aldex = dda_aldex_prevalente$objetos_aldex), ncol= 2))
do.call(grid.arrange, c(volcanos_aldex_flr(objetos_aldex = dda_aldex_rara$objetos_aldex), ncol= 2))
```

<https://cran.r-project.org/web/packages/forestploter/vignettes/forestploter-intro.html>

```{r, fig.width=8, fig.height=4}

lista <- list()
for (i in dda_aldex_prevalente$taxa_relevante$variable) {
    
    microorganismos <- dda_aldex_prevalente$taxa_relevante %>%
        dplyr::filter(variable == i) %>%
        pull(code)
    
    lista[[i]] <- microorganismos
    
}

upset(fromList(lista), 
      order.by = "freq",
      sets.x.label = "Taxonomy by variable",
      mainbar.y.label = "Taxonomy",
      point.size = 3.5, 
      line.size = 1.5,
      text.scale = c(1.3, 1.3, 1, 1, 1.5, 1.3),

)

```

### ANCOM-BC2

<https://www.bioconductor.org/packages/release/bioc/vignettes/ANCOMBC/inst/doc/ANCOMBC2.html>

#### Species

```{r, warning=FALSE, message=FALSE}
# Agglomerate by genus and subset by prevalence
tse_prevalente <- subsetByPrevalent(tse_lp, rank = "Species", prevalence = 0.1)
tse_rara <- subsetByRare(tse_lp, rank = "Species", prevalence = 0.23)

# Transform count assay to relative abundances
tse_prevalente <- transformAssay(tse_prevalente, assay.type = "counts", method = "relabundance")
tse_rara <- transformAssay(tse_rara, assay.type = "counts", method = "relabundance")
```

```{r, eval=FALSE}
dda_ancombc_prevalente <- ancombc2_flr(objeto_tse = tse_prevalente, prevalence = "prevalente", taxa_relevante = taxa_relevante_species)

dda_ancombc_rara <- ancombc2_flr(objeto_tse = tse_rara, prevalence = "rara", taxa_relevante = taxa_relevante_species)

saveRDS(dda_ancombc_prevalente, "../RDSs/dda_ancombc_prevalente.RDS")
saveRDS(dda_ancombc_rara, "../RDSs/dda_ancombc_rara.RDS")

```

```{r}
dda_ancombc_prevalente <- readRDS("../RDSs/dda_ancombc_prevalente.RDS")
dda_ancombc_rara <- readRDS("../RDSs/dda_ancombc_rara.RDS")
```

```{r, fig.width=10, fig.height=7, warning=FALSE}
do.call(grid.arrange, c(volcanos_ancombc_flr(objetos_ancombc = dda_ancombc_prevalente$objetos_ancombc), ncol= 2))
do.call(grid.arrange, c(volcanos_ancombc_flr(objetos_ancombc = dda_ancombc_rara$objetos_ancombc), ncol= 2))
```

<https://cran.r-project.org/web/packages/forestploter/vignettes/forestploter-intro.html>

```{r, fig.width=8, fig.height=4}

lista <- list()
for (i in dda_ancombc_prevalente$taxa_relevante$variable) {
    
    microorganismos <- dda_ancombc_prevalente$taxa_relevante %>%
        dplyr::filter(variable == i) %>%
        pull(taxon)
    
    lista[[i]] <- microorganismos
    
}

upset(fromList(lista), 
      order.by = "freq",
      sets.x.label = "Taxonomy by variable",
      mainbar.y.label = "Taxonomy",
      point.size = 3.5, 
      line.size = 1.5,
      text.scale = c(1.3, 1.3, 1, 1, 1.5, 1.3),

)

```

```{r, fig.width=8, fig.height=4}

lista <- list()
for (i in dda_ancombc_rara$taxa_relevante$variable) {
    
    microorganismos <- dda_ancombc_rara$taxa_relevante %>%
        dplyr::filter(variable == i) %>%
        pull(taxon)
    
    lista[[i]] <- microorganismos
    
}

upset(fromList(lista), 
      order.by = "freq",
      sets.x.label = "Taxonomy by variable",
      mainbar.y.label = "Taxonomy",
      point.size = 3.5, 
      line.size = 1.5,
      text.scale = c(1.3, 1.3, 1, 1, 1.5, 1.3),

)

```

#### Genus

```{r, warning=FALSE, message=FALSE}
# Agglomerate by genus and subset by prevalence
tse_prevalente <- subsetByPrevalent(tse_lp, rank = "Genus", prevalence = 0.1)
tse_rara <- subsetByRare(tse_lp, rank = "Genus", prevalence = 0.23)

# Transform count assay to relative abundances
tse_prevalente <- transformAssay(tse_prevalente, assay.type = "counts", method = "relabundance")
tse_rara <- transformAssay(tse_rara, assay.type = "counts", method = "relabundance")
```

```{r, eval=FALSE}
dda_ancombc_prevalente <- ancombc2_flr(objeto_tse = tse_prevalente, prevalence = "prevalente", taxa_relevante = taxa_relevante_species)

dda_ancombc_rara <- ancombc2_flr(objeto_tse = tse_rara, prevalence = "rara", taxa_relevante = taxa_relevante_species)

saveRDS(dda_ancombc_prevalente, "../RDSs/dda_ancombc_prevalente_genus.RDS")
saveRDS(dda_ancombc_rara, "../RDSs/dda_ancombc_rara_genus.RDS")

```

```{r}
dda_ancombc_prevalente <- readRDS("../RDSs/dda_ancombc_prevalente_genus.RDS")
dda_ancombc_rara <- readRDS("../RDSs/dda_ancombc_rara_genus.RDS")
```

```{r, fig.width=10, fig.height=7, warning=FALSE}
do.call(grid.arrange, c(volcanos_ancombc_flr(objetos_ancombc = dda_ancombc_prevalente$objetos_ancombc), ncol= 2))
do.call(grid.arrange, c(volcanos_ancombc_flr(objetos_ancombc = dda_ancombc_rara$objetos_ancombc), ncol= 2))
```

<https://cran.r-project.org/web/packages/forestploter/vignettes/forestploter-intro.html>

```{r, fig.width=8, fig.height=4}

lista <- list()
for (i in dda_ancombc_prevalente$taxa_relevante$variable) {
    
    microorganismos <- dda_ancombc_prevalente$taxa_relevante %>%
        dplyr::filter(variable == i) %>%
        pull(taxon)
    
    lista[[i]] <- microorganismos
    
}

upset(fromList(lista), 
      order.by = "freq",
      sets.x.label = "Taxonomy by variable",
      mainbar.y.label = "Taxonomy",
      point.size = 3.5, 
      line.size = 1.5,
      text.scale = c(1.3, 1.3, 1, 1, 1.5, 1.3),

)

```

```{r, fig.width=8, fig.height=4}

lista <- list()
for (i in dda_ancombc_rara$taxa_relevante$variable) {
    
    microorganismos <- dda_ancombc_rara$taxa_relevante %>%
        dplyr::filter(variable == i) %>%
        pull(taxon)
    
    lista[[i]] <- microorganismos
    
}

upset(fromList(lista), 
      order.by = "freq",
      sets.x.label = "Taxonomy by variable",
      mainbar.y.label = "Taxonomy",
      point.size = 3.5, 
      line.size = 1.5,
      text.scale = c(1.3, 1.3, 1, 1, 1.5, 1.3),

)

```

# Association between taxa








